---
title: "Exploring U.S. Supreme Court Decisions"
author: "Abby Bergman"
output: github_document
---

## Get the data

```{r, echo = TRUE}
# load useful packages
library(tidyverse)

# load data
scdbv_mod <- read_csv("data/SCDB_2018_01_justiceCentered_Citation.csv", guess_max = 100000)
scdbv_leg <- read_csv("data/SCDB_Legacy_04_justiceCentered_Citation.csv", guess_max = 100000)
scdbv_mod
scdbv_leg
```

## Combine the datasets

```{r}
scdbv <- bind_rows(scdbv_leg, scdbv_mod)

```

## Recode variables as you find necessary

```{r}
scdbv <- scdbv %>%
mutate(majority = majority -1, datadecisionDirection = ifelse(decisionDirection == 3, NA, decisionDirection -1), justiceName = as.factor(justiceName),
         chief = as.factor(chief))

```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}

scdbv2 <- scdbv %>%
  mutate(margin = majVotes-minVotes) %>% #create margin column
  select(caseId, term, majVotes, minVotes, margin) %>% #select only the terms I want
  unique() %>%
  group_by(term)

scdbv2%>%
  count(margin) %>% 
  mutate(percent = n / sum(n))%>%
  filter(margin == 1) %>%
  ggplot(aes(term, percent))+ geom_line() + labs(title = "Percentage of Votes with a One Vote Margon", x = "Term", y = "Percentage") + scale_y_continuous(labels = scales::percent)

```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}
  
scdbv %>%
  filter(justiceName == "AScalia")%>%
  count(term, majority) %>%
  na.omit() %>%
  group_by(term) %>%
  summarise(all_n = sum(n), 
            all_maj = sum(n[majority==1])) %>%
  mutate(n_pct = all_maj/all_n) %>%
  ggplot(aes(term, n_pct)) + geom_line() + scale_y_continuous(labels=scales::percent, limits=c(0,NA)) +
  labs(title= "Percent of Time Scalia was in the Majority", x = "Term", y= "Percent of Total Decisions")
  

```

## Create a graph similar to above that compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}
scdbv3 <- scdbv %>%
  filter(justiceName == "AScalia")%>%
  mutate(unani = ifelse(minVotes == 0,1,0)) %>%
  count(term, majority, unani) %>%
  na.omit() %>%
  group_by(term) %>%
  summarise(all_n = sum(n), all_maj = sum(n[majority==1]), nonUnani_maj = sum(n[majority==1 & unani==0 ]), nonUnani_n = sum(n[unani==0])) %>%
  gather(key = key, value = value, -term) %>%
  separate(key, into = c("type", "var"))%>%
  spread(var, value)%>%
  mutate(n_pct = maj/n, 
         type = factor(type, levels = c("all", "nonUnani"), labels = c("All Decisions", "Non-Unanimous Decisions")))
  
ggplot(scdbv3, aes(term, n_pct, color=type))+
  geom_line()+
  scale_y_continuous(labels = scales::percent, limits=c(0,NA))+ 
  labs(titles = "Percent of Time Scalia was in the Majority", x="Term", y = "Percent of Total Decisions") + 
  theme(legend.position="bottom")

```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}
#combine data sets
question4 <- bind_rows(scdbv_leg, scdbv_mod)

question4 %>%
  count(term, decisionDirection)%>%
  group_by(term) %>%
  na.omit %>%
  summarize(all_n =sum(n),
            conservative = sum(n[decisionDirection == 1])) %>%
  ggplot(aes(term,conservative/all_n)) + 
  geom_line()+
  labs(title= "Percent of US Supreme Court Cases Won in a Conservative Direction", x = "Term", y = "Percent of Cases") +
  scale_y_continuous(labels = scales::percent, limits=c(0,NA))

scdbv4 <- scdbv %>%
  select(term, caseId, partyWinning, decisionDirection) %>% #select only the terms I want
  unique() %>%
  group_by(term)

```

## In each term, how many of the term's published decisions were announced in a given month?

```{r}

scdbv %>%
  unique()%>%
  select(term, dateDecision) %>%
  mutate(Month = dateDecision) %>%
  separate(Month, into = c("month", "day", "year")) %>%
  select(term, dateDecision, month) %>%
  mutate(month =factor(month, levels = c(1:12), labels = c(month.abb)))%>%
  group_by(term)%>%
 count(term, month) %>%
  ggplot(aes(month, n))+ 
  geom_boxplot() +
  coord_flip() 


scdbv%>%
  unique()%>%
  select(term, dateDecision)%>%
  


library(lubridate)
scdbv %>%
  mutate(decision_month=month(mdy(dateDecision)), factor(decision_month, labels = month.abb))%>%
  group_by(decision_month)%>%
  summarize(avg_decision = n()) %>%
mutate(decision_month =factor(decision_month, levels = c(1:12), labels = c(month.abb)))%>%
 
  count(decision_month) %>%
  ggplot(aes(decision_month, n))+ 
  geom_boxplot() +
  coord_flip() 


test5 <- scdbv%>%
  mutate(decision_month = month(mdy(dateDecision))) %>%
  mutate(decision_month = factor(decision_month, labels = month.name))

test6 <- test5 %>%
  select(term, decision_month, caseIssuesId)

test7 <- test6[!duplicated(test6$caseIssuesId),] %>%
  group_by(term, decision_month) %>%
  summarise(total_count = n())

ggplot(test7, aes(decision_month, total_count)) +
  geom_boxplot()+
  coord_flip() +
  labs(title = "Number of Published Decisions Per Month, by Term", x = "Term Month", y = "Number of Published Decisions in a Term-Month")



test7
```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.

```{r}
scdbv6 <- scdbv %>%
  select(term, caseId, partyWinning, decisionDirection, chief) %>% #select only the terms I want
  unique() %>%
  group_by(term, chief)

scdbv6%>%
  count(decisionDirection) %>% 
  mutate(percent = n / sum(n))%>%
  filter(decisionDirection == 1)  %>%
  group_by(chief) %>%
  ggplot(aes(x = term, y = percent))+
  geom_path()+ 
  facet_wrap(~chief) + 
  labs(title = "Percent of Supreme Court Cases Decided in a Conservative Direction", subtitle = "Showing Chief Justice for a Given Term", y = "Percent of Cases", x = "Term") +
  scale_y_continuous(labels = scales::percent, limits=c(0,NA)) 

```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


