---
title: "Exploring U.S. Supreme Court Decisions"
author: "Abby Bergman"
output: github_document
---

## Get the data

```{r, echo = TRUE}
# load useful packages
library(tidyverse)

# load data
scdbv_mod <- read_csv("data/SCDB_2018_01_justiceCentered_Citation.csv", guess_max = 100000)
scdbv_leg <- read_csv("data/SCDB_Legacy_04_justiceCentered_Citation.csv", guess_max = 100000)
scdbv_mod
scdbv_leg
```

## Combine the datasets

```{r}
scdbv <- bind_rows(scdbv_leg, scdbv_mod)

```

## Recode variables as you find necessary

```{r}

```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}
scdbv2 <- scdbv %>%
  mutate(margin = majVotes-minVotes) %>% #create margin column
  select(caseId, term, majVotes, minVotes, margin) %>% #select only the terms I want
  unique() %>%
  group_by(term)


#this is wrong but im going to leave for now
# what im missing is some way to summarize each term
scdbv2%>%
  count(margin) %>% 
  mutate(percent = n / sum(n))%>%
  filter(margin == 1) %>%
  ggplot(aes(term, percent))+ geom_line() + labs(title = "Percentage of Votes with a One Vote Margon", x = "Term", y = "Percentage") + scale_y_continuous(labels = scales::percent)

```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}
scdbv %>%
  select(caseId, term, majVotes, minVotes, justice, majority, justiceName) %>%
  filter(term > 1985)
  

```

## Create a graph similar to above that compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}

```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}

```

## In each term, how many of the term's published decisions were announced in a given month?

```{r}

```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.

```{r}

```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


